NumPy---Array Programming in Python

### 2020-02-06 body.tex
Redraw Fig. 1 in Inkscape
Refine conceps sketch; update captions

Two Python array packages existed before NumPy.
The Numeric package began in the mid-1990s and provided an array object and array-aware functions
in Python, written in C, and linking to standard fast implementations of linear
algebra \cite{Numericmanual}.
To handle large astronomical images coming from the Hubble Space Telescope,
a reimplementation of Numeric, called NumArray, added
support for memory-mapped arrays and mixed data type records. 
%Around 2000, the Space Telescope Science Institute (STScI) software group wrote
%a reimplementation of Numeric, called NumArray, to support their work
%on large memory-mapped arrays and arrays of mixed data type
%records \cite{greenfield2003numarray}.
While NumArray was highly compatible with Numeric, the two packages
had enough differences that it divided the community, until
2005, when NumPy emerged as a ``best of both worlds'' unification of Numeric
and NumArray \cite{oliphant2006guide}.

% KJM: I moved this here to get the placement right 
\begin{figure*}[h]
  \centering
  \includegraphics[width=\textwidth]{static/sketches/array-concepts}   
  \caption{\textbf{Fundamental Array Concepts.}
    \textbf{a,} The NumPy array data structure and its associated metadata fields.
    \textbf{b,} Indexing an array with various types of arguments.
    \textbf{c,} Indexing with arrays, which broadcast the indexing arguments before performing the lookup.
    \textbf{d,} Vectorization efficiently applies operations to groups of elements
    \textbf{e,} Broadcasting in scalar addition, and in the division of two-dimensional arrays.
    \textbf{f,} Reduction operations act along one or more axes. In this
    example, a three-dimensional array is shown to be summed along various single
    axes to produce two-dimensional results, or along two axes consecutively to
    produce a one-dimensional result.
    \textbf{g,} Example NumPy code, illustrating some of these concepts.
   }
  \label{fig:array-concepts}
\end{figure*}

>>>>

Two Python array packages existed before NumPy.
The Numeric package began in the mid-1990s and provided an array object and array-aware functions
in Python, written in C, and linking to standard fast implementations of linear
algebra \cite{Numericmanual}.
To handle large astronomical images coming from the Hubble Space Telescope,
a reimplementation of Numeric, called NumArray, added
support for memory-mapped arrays and mixed data type records. 
While NumArray was highly compatible with Numeric, the two packages
had enough differences that it divided the community, until
2005, when NumPy emerged as a ``best of both worlds'' unification of Numeric
and NumArray \cite{oliphant2006guide}.

% KJM: I moved this here to get the placement right
% FIXME SJW: let's review caption (d); I don't think it's crystal clear what
% we're trying to show there; double check caption (g)
\begin{figure*}[h]
  \centering
  \includegraphics[width=\textwidth]{static/sketches/concepts}
  \caption{\textbf{Fundamental Array Concepts.}
    \textbf{a,} The NumPy array data structure and its associated metadata fields.
    \textbf{b,} Indexing an array with slices and steps. These
  operations return a \emph{view} of the original data.
    \textbf{c,} Indexing an array with masks, scalar coordinates, or
  other arrays, so that it returns a copy of the original data. In the
  bottom example, an array is indexed with other arrays; this broadcasts the indexing arguments before performing the lookup.
    \textbf{d,} Vectorization efficiently applies operations to groups
  of elements.
    \textbf{e,} Broadcasting in the multiplication of two-dimensional arrays.
    \textbf{f,} Reduction operations act along one or more axes. In this
    example, an array is be summed along various axes to produce a vector, or along two axes consecutively to
    produce a scalar.
    \textbf{g,} Example NumPy code, illustrating some of these concepts.
   }
  \label{fig:array-concepts}
\end{figure*}
